---
title: "Simulation 1"
author: "Lily Andrews"
date: "2023-11-28"
output: github_document
---
Install packages
```{r}
library(tibble)
library(simulateGP)
library(dplyr)
```

Simulation 1

Data generating model:
```{r}
rmarkdown::render(".R/sims_functions.Rmd")
dat <- dgmodel(
  nid=100000, #number of individuals
  nsnp=99, #number of snps
  gc_maf=runif(99, 0.05, 0.95), #no rare variants included
  b_gl=rnorm(99), 
  rsq_prs=0.06, #h-squared value
  rsq_z=0.19, #to total 0.25 of total H-squared
  d_prev=5/100000, 
  b_lr0=sqrt(0.1), 
  b_c0l=sqrt(0.01), 
  gr_maf=0.5, 
  rsq_gr0=0.1,   # this is a pqtl - often large effect size
  rsq_gc0=0.1,   # this is a pqtl - often large effect size 
  b_u1r1=0.1, 
  b_u1l=0.1, 
  b_u2c1=0.1, 
  b_u2l=0.1, 
  b_dr1=0.1, 
  b_dc1=0.1, 
  b_c0c1=0.1, 
  b_r0r1=0.1, 
  b_u2c0=0.1, 
  b_gcc0=0.1, 
  b_lc0=0.1, 
  b_u1r0=0.1, 
  b_u1d=0.1, 
  b_u2d=0.1,
  b_gcc1=0.1)

dat$phen
```
Variance and means of variables from data generating model and correlation between PRS and disease liability
```{r}
dgmodel_check(dat)
```
```{r}
dgmodel_analysis(dat, ncase=1000, ncontrol=1000, protein_gwas=1000)
```

##to trial
#prevalence of disease in population - e.g. if no disease in population can it still predict?
Scenerio - Alter heritability explained
```{r}
set.seed(100)
reps = 1000
scenerio <- 1
dat <- dgmodel(
  nid=100000, #number of individuals
  nsnp=99, #number of snps
  gc_maf=runif(99, 0.05, 0.95), #no rare variants included
  b_gl=rnorm(99), 
  rsq_prs=0.06, 
  rsq_zl=0.19, 
  d_prev=5/1000, 
  b_lr0=sqrt(0.1), 
  b_c0l=sqrt(0.01), 
  gr_maf=0.5, 
  rsq_gr0=0.1,   # this is a pqtl - often large effect size
  rsq_gc0=0.1,   # this is a pqtl - often large effect size 
  b_u1r1=0.1, 
  b_u1l=0.1, 
  b_u2c1=0.1, 
  b_u2l=0.1, 
  b_dr1=0.1, 
  b_dc1=0.1, 
  b_c0c1=0.1, 
  b_r0r1=0.1, 
  b_u2c0=0.1, 
  b_gcc0=0.1, 
  b_lc0=0.1, 
  b_u1r0=0.1, 
  b_u1d=0.1, 
  b_u2d=0.1)
```

Scenerio - Alteration of causal SNPs
```{r}
set.seed(13)
reps = 1000
scenerio <- 2
dat <- dgmodel(
  nid=100000, #number of individuals
  nsnp=99, #number of snps
  gc_maf=runif(99, 0.05, 0.95), #no rare variants included
  b_gl=rnorm(99), 
  rsq_prs=0.06, 
  rsq_zl=0.19, 
  d_prev=5/1000, 
  b_lr0=sqrt(0.1), 
  b_c0l=sqrt(0.01), 
  gr_maf=0.5, 
  rsq_gr0=0.1,   # this is a pqtl - often large effect size
  rsq_gc0=0.1,   # this is a pqtl - often large effect size 
  b_u1r1=0.1, 
  b_u1l=0.1, 
  b_u2c1=0.1, 
  b_u2l=0.1, 
  b_dr1=0.1, 
  b_dc1=0.1, 
  b_c0c1=0.1, 
  b_r0r1=0.1, 
  b_u2c0=0.1, 
  b_gcc0=0.1, 
  b_lc0=0.1, 
  b_u1r0=0.1, 
  b_u1d=0.1, 
  b_u2d=0.1)
```
Apply this to pQTL data (deCODE)
```{r}
#first example
dat <- dgmodel(
  nid=35559, #number of individuals
  nsnp=99, #number of snps
  gc_maf=runif(99, 0.05, 0.95), #no rare variants included
  b_gl=rnorm(99), 
  rsq_prs=0.06, #h-squared value https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4667278/
  rsq_zl=0.19, #to total 0.25 of total H-squared https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4667278/
  d_prev=5/1000, #~5% http://rarecarenet.istitutotumori.mi.it/analysis.php
  b_lr0=sqrt(0.1), 
  b_c0l=sqrt(0.01), 
  gr_maf=0.5, 
  rsq_gr0=0.1,   # this is a pqtl - often large effect size
  rsq_gc0=0.1,   # this is a pqtl - often large effect size 
  b_u1r1=0.1, 
  b_u1l=0.1, 
  b_u2c1=0.1, 
  b_u2l=0.1, 
  b_dr1=0.1, 
  b_dc1=0.1, 
  b_c0c1=0.1, 
  b_r0r1=0.1, 
  b_u2c0=0.1, 
  b_gcc0=0.1, 
  b_lc0=0.1, 
  b_u1r0=0.1, 
  b_u1d=0.1, 
  b_u2d=0.1)

system.time(output <- lapply(replication, dgmodel)) 
```
Apply this to sun
```{r}
dgmodel <- function(nid=3301, 
  nsnp=99, 
  gc_maf=runif(99, 0.05, 0.95), #99 SNPs and no rare variants
  b_gl=rnorm(99), 
  rsq_prs=0.06, #h-squared value https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4667278/
  rsq_zl=0.19, #to total 0.25 of total H-squared https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4667278/
  d_prev=5/1000, #~5% http://rarecarenet.istitutotumori.mi.it/analysis.php
  b_lr0=sqrt(0.1), 
  b_c0l=sqrt(0.01), 
  gr_maf=0.5, 
  rsq_gr0=0.1,   
  rsq_gc0=0.1,   
  b_u1r1=0.1, 
  b_u1l=0.1, 
  b_u2c1=0.1, 
  b_u2l=0.1, 
  b_dr1=0.1, 
  b_dc1=0.1, 
  b_c0c1=0.1, 
  b_r0r1=0.1, 
  b_u2c0=0.1, 
  b_gcc0=0.1, 
  b_lc0=0.1, 
  b_u1r0=0.1, 
  b_u1d=0.1, 
  b_u2d=0.1)
{
  u1 <- rnorm(nid) 
  u2 <- rnorm(nid) 
  gc <- make_geno(nid, nsnp, gc_maf) 
  gr <- rbinom(nid, 2, gr_maf) 
  c0 <- scale(gc[,1]) * b_gcc0 + u2 * b_u2c0 + rnorm(nid, sd=sqrt(1-rsq_gc0-b_lc0^2-b_u2c0^2)) 
  rsq_prs <- rsq_prs - (0.1)^2 
  prs <- scale(gc %*% b_gl) 
  prs_w <- scale(gc[,-1] %*% b_gl[-1]) 
  z <- rnorm(nid) 
  l <- prs_w * sqrt(rsq_prs) + z * sqrt(rsq_zl) + c0 * b_c0l + u1 * b_u1l + u2 * b_u2l 
  r0 <- scale(gr) * sqrt(rsq_gr0) + l * b_lr0 + u1 * b_u1r0 + rnorm(nid, sd=sqrt(1-rsq_gr0-b_lr0^2-b_u1r0^2)) 
  prob_l <- simulateGP::gx_to_gp(gx=scale(l), h2x=rsq_prs + rsq_zl + b_c0l^2 + b_u1d^2 + b_u2d^2, prev = d_prev) 
  d <- rbinom(nid, 1, prob_l)
  d <- abs(d-1) 
  c1 <- scale(gc[,1]) * 0.1 + u2 * b_u2c1 + c0 *b_c0c1 + d * b_dc1 + rnorm(nid, sd=sqrt(1-(0.1)^2 - b_u2c1^2- b_c0c1^2)) 
  r1 <- scale(gr) * sqrt(rsq_gr0) + r0 * b_r0r1 + u1 * b_u1r1 + d * b_dr1 + rnorm(nid, sd=sqrt(1-rsq_gc0-b_lr0^2-b_u1r1^2)) 
  phen <- tibble(
    u1, u2, r0, r1, c0, c1, prs, prs_w, l, prob_l, d
  )
  return(list(geno=gc, phen=phen)) 
}
system.time(results <- lapply(replication, dgmodel)) #time without paralleisation
system.time(results <- mclapply(replication, dgmodel, mc.cores = numCores)) #time with addition of parallelisation
```

